Palomar 5 Gaps

### 2024-09-30 paper/main.tex
edited the overview

  \subsection*{Overview}
  This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their corresponding uncertainties and then repeat the experiment multiple times. We choose to perform the simulation 50 times to balance computational cost with a high enough number to allow us to explore a range of possible outcomes. The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere allowed in their phase space given their initial energy and angular momenta. We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps. Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.

  For each completed simulation, we follow the scheme laid out in the methods. I.e. we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions, with two different cases. The \textit{Full} simulation that includes the influence of all globualr clusters on the star particles representing Palomar 5 and its stream, as well as the \textit{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5. We do this for we attempt to quantify the number of gaps from globular clusters in the limit of maximum information. I.e. we are not limited by contamination of field stars, and since we are computing our own simulations, we know what a gapless stream should look like.

  This comparative analysis is shown in Fig.~\refP{fig:profiles}. We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit. As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations. We find that many streams as visible by eye. However, some subtle under densities only became apparent when comparing the 1D density profiles. All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.

  Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps. This method used the orbit of Palomar 5 as a proxy for the stream. The gravitational force from all other globular clusters as computed at segments of the orbit about the position of Palomar 5 for the entire simulation. The top 5 strongest flybys were recorded to identify which cluster was responsible. This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream. A density threshold was imparted to quantify the length of the stream.  The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible.

  Fig.~\ref{fig:force-on-orbit} contains some interesting information. Notice the periodic ribbions of force in the $(t,\tau)$ plane. This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system. Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map. Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap. Notice how NGC5272, which was labeled with 0 to indicate that it has the greatest local maxima does not have a gap. This is manifold, for instance, the force needs to be modulated with the time of impact because the  important quantity that determins gap creation.  Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates, so peaks upon the orbit are good proxy's for the stream but are not definitive. We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.

 For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}. Note that the gap from NGC~7078 was \textit{hit twice}.

>>>>

  \subsection*{Overview}
  This experiment follows a monte-carlo scheme, where we sample the initial conditions of the globular clusters' masses and kinematics based on their reported observational uncertainties and repeat the experiment multiple times. We choose to perform the simulation 50 times to balance computational and exploring a range of possible outcomes. The uncertainties on the globular clusters are large enough that over the course of the 5 Gyr integration period, they could be anywhere in their phase space given their initial energy and angular momentum. We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps. Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.

  For each completed simulation, we follow the scheme laid out in the methods. I.e. we simulate the creation of the Palomar 5 stream for 50 different samplings of the initial conditions in two different environments. The first are the \texttt{vanilla} simulations in which the star-particles feel only the galaxy and Palomar 5. On the other hand, the \texttt{full} simulations includes the influence of all globualr clusters on the star-particles.

  By performing a comparative analysis, we are exploring the phenomenon of the influence from all globular clusters on stellar streams in a regieme of far more complete information respect to reality. In this experiment, we do not deal with differentiating stream and field stars, and above all else in reality we do not have a reference reality without the effect of other globular clusters as a point of comparison. Additionally, we avoid the difficulty of attempting to define a gap in its own right. Erkal et al 2015 was able to do this pretty well by studying two parameters: the width of the gaps and the ratio of the minimum density to the density of the gaps' shoulders. However, this becomes impossible when departing from the idealized case of uniformly dense \& one-dimensional streams on circular orbits. Sanders et al 2016 extended this in action-angle space which allowed for dropping the constraint of circular orbits. However, I need to revist the article again because I believe they also explored non uniformly dense streams. Nonetheless, we avoid this all together by doing our comparative analysis. Since we are numerical modellers, we can play god and run nearly the exact same experiment by including or not including the globular clusters. Therefore, it's much easier to identify what comes from the clusters or not.

  This comparative analysis is shown in Fig.~\ref{fig:profiles}. We visually compare the 2D density map in tail coordinates as well as their marginalized 1D density profiles along the length of Palomar 5's orbit. As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations. We find that many gaps are visible by eye. However, some subtle under densities only became apparent when comparing the 1D density profiles. All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}.

  Subsequently, we followed the analysis presented in section~\ref{sec:SuspectToCulprit} to identify which globualr clusters were responsible for the gaps. This method used the orbit of Palomar 5 as a proxy for the stream. This is contrary to the typical case where one uses the stream to approximate the orbit--but $\approx$ works both ways. The gravitational force from all other globular clusters was computed at segments of the orbit about the position of Palomar 5 for the entire simulation. The top 5 strongest flybys were identified and the corresponding perturbers are labeled as \textit{suspects}. This is shown in Fig.~\ref{fig:force-on-orbit}, where the top panel shows the force from the globular clusters on the orbit, while the bottom panel shows the density of the stream. A density threshold was imparted to quantify the length of the stream. The location and time of impact was cross referenced to the location of the gaps to identify which cluster was responsible. I.e. if the gap was observed on the leading tail, we searched for suspects with $\tau>0$.

  The top panel of Fig.~\ref{fig:force-on-orbit} contains some interesting information. Notice the periodic ribbions of force in the $(t,\tau)$ plane. This is due to pericenter passages where Palomar 5 is getting closer to the center of mass of the globular cluster system. Additionally, for the impacts of NGC104 and NGC7078, wakes can be observed in the density map---evidence that signifies their guilt. Another important aspect, is that the strongest peak in gravitational force does not necessarily create a gap. Notice how NGC~5272, which was labeled with 0 to indicate that it has the greatest force of all local maxima, does not provoke a gap. The reason for this is manifold. First, remember that the important quantity is \textit{the change in momentum}. Thus, a strong force applied for a brief peroid of time may have no effect on the stream!  Additionally, there is an offset of about 200~pc between the stream at the orbit, as seen when viewing the stream in tail coordinates. This demonstrates that the orbit are good proxy's for the stream but are not a one for one substitute. We found that the top five greatest impacts accounted for all gaps, except for Sampling~014 as shown in Fig.~\ref{fig:gallery3}, whose gap from NGC~6584 corresponded to the 7th peak.

 For some cases, we performed a double check by launching extra simulations only including one suspect at a time to confirm that this suspect indeed caused a gap in question, an example is shown in Fig.~\ref{fig:decomposition}. Note that the gap from NGC~7078 was \textit{hit} by NGC~7078 \textit{twice}.
